---
title: "Installing skeleSim"
author: "A Strand"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Installing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette focuses on getting everything together to get this package to run.

##CRAN Dependencies

There are several packages located on CRAN that skeleSim depends upon.  They all currently install automatically during the installation process.  

##Non-CRAN dependencies

Currently, skeleSim implements coalescent simulations in `fastsimcoal2` and forward-time in `rmetasim`. The `rmetasim` dependency is handled via normal package installation.

You must install fastsimcoal2 on your computer yourself.  The executable for your operating system (Windows, MacOSX, linux) is available [here](http://cmpg.unibe.ch/software/fastsimcoal2).

**The trick is to make sure that the folder/directory in which the fastsimcoal executable resides is in the search path for your operating system**

Helpful instructions for editing the PATH environmental variable for Windows can be found [here](https://www.java.com/en/download/help/path.xml). On Mac OS or LINUX/UNIX systems, the executable should be placed in the _/usr/local/bin_ folder which is usually a default in the PATH.

If the executable is in your path, skeleSimGUI() looks for it (and other versions if present) and allows you to select among versions on the "FastSimCoal params" tab.

##Installing from github
The most up to date version (new bugs and all) of skeleSim is always going to be on github.  This is also a place to ask specific questions about skeleSim.

First install the great 'devtools' package. Then,

```{r,eval=FALSE}
library(devtools)
install_github("christianparobek/skeleSim", build_vignettes = TRUE, force = TRUE)
```

Once the above has been completed the shiny interface can be run:

```{r,eval=FALSE}
library(skeleSim)
skeleSimGUI()
```
